FAIR Theory: Applying Open Science Principles to the Construction and Iterative Improvement of Scientific Theories

### 2024-12-19 manuscript.Rmd
update 2024-12-19 11:45:49.650496

If we consider the current state of Reusability in psychological theory, there appears to be a norm *against* theory reuse:
*"[Theories are] like toothbrushes — no self-respecting person wants to use anyone else's"* [@mischelToothbrushProblem2008].
This norm impedes scientific progress.
Cumulative knowledge acquisition requires reusable theories that are continuously updated based on insights from new data [@degrootMethodologieGrondslagenVan1961].
<!-- Theories are made reusable by thorough documentation, -->
<!-- appropriate licensing,  -->
<!-- and detailed provenance (e.g., through version control, cross-referencing, and semantic versioning). -->
In our workshops on FAIR theory, we similarly notice reluctance to the notion of reusing and adapting theories,
reflected in questions such as "who owns theory",
and "who determines how a theory may be reused or changed"?
These questions imply a norm against modifying theory without consent from the author reminiscent of the aforementioned problem of dependency on the author.

>>>>

If we consider the current state of Reusability in psychological theory, there appears to be a norm *against* theory reuse:
*"[Theories are] like toothbrushes — no self-respecting person wants to use anyone else's"* [@mischelToothbrushProblem2008].
This norm impedes scientific progress.
Cumulative knowledge acquisition requires reusable theories that are continuously updated based on insights from new data [@degrootMethodologieGrondslagenVan1961].
In our workshops on FAIR theory, we similarly notice reluctance to the notion of reusing and adapting theories,
reflected in questions such as "who owns theory",
and "who determines how a theory may be reused or changed"?
These questions imply a norm against modifying theory without consent from the author reminiscent of the aforementioned problem of dependency on the author.
